### 2017-06-10 report/paper.tex
report update

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% University Assignment Title Page 
% LaTeX Template
% Version 1.0 (27/12/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% WikiBooks (http://en.wikibooks.org/wiki/LaTeX/Title_Creation)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
% 
% Instructions for using this template:
% This title page is capable of being compiled as is. This is not useful for 
% including it in another document. To do this, you have two options: 
%
% 1) Copy/paste everything between \begin{document} and \end{document} 
% starting at \begin{titlepage} and paste this into another LaTeX file where you 
% want your title page.
% OR
% 2) Remove everything outside the \begin{titlepage} and \end{titlepage} and 
% move this file to the same directory as the LaTeX file you wish to add it to. 
% Then add \input{./title_page_1.tex} to your LaTeX file where you want your
% title page.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\title{Title page with logo}
%----------------------------------------------------------------------------------------
% PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[paper=a4, fontsize=11pt]{scrartcl}
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}

\begin{document}

\begin{titlepage}

\newcommand{\HRule}{\rule{\linewidth}{0.5mm}} % Defines a new command for the horizontal lines, change thickness here

\center % Center everything on the page

%----------------------------------------------------------------------------------------
% HEADING SECTIONS
%----------------------------------------------------------------------------------------

\textsc{\LARGE EURECOM}\\[1.5cm] % Name of your university/college
%\textsc{\Large Major Heading}\\[0.5cm] % Major heading such as course name
%\textsc{\large Minor Heading}\\[0.5cm] % Minor heading such as course title

%----------------------------------------------------------------------------------------
% TITLE SECTION
%----------------------------------------------------------------------------------------

\HRule \\[0.4cm]
{ \huge \bfseries Music Recommendation when Exploring a City}\\[0.4cm] % Title of your document
\HRule \\[1.5cm]

%----------------------------------------------------------------------------------------
% AUTHOR SECTION
%----------------------------------------------------------------------------------------

\begin{minipage}{0.4\textwidth}
\begin{flushleft} \large
\emph{Authors:}\\
Lorenzo \textsc{Canale} % Your name
\newline
Fabio \textsc{Ellena} % Your name
\end{flushleft}
\end{minipage}
~
\begin{minipage}{0.4\textwidth}
\begin{flushright} \large
\emph{Supervisor:} \\
Pasquale \textsc{Lisena} % Supervisor's Name
\end{flushright}
\end{minipage}\\[2cm]

% If you don't want a supervisor, uncomment the two lines below and remove the section above
%\Large \emph{Author:}\\
%John \textsc{Smith}\\[3cm] % Your name

%----------------------------------------------------------------------------------------
% DATE SECTION
%----------------------------------------------------------------------------------------

%{\large \today}\\[2cm] % Date, change the \today to a set date if you want to be precise

%----------------------------------------------------------------------------------------
% LOGO SECTION
%----------------------------------------------------------------------------------------

%\includegraphics{logo.png}\\[1cm] % Include a department/university logo - this will require the graphicx package

%----------------------------------------------------------------------------------------

\vfill % Fill the rest of the page with whitespace

\end{titlepage}
\tableofcontents
\newpage
\begin{abstract}
Abstract: In this project we show a new approach for finding artists linked to a place of interest (POI). We propose a semantic-based framework built upon the DBpedia, 3cixty, and Doremus. Furthermore we present a graph exploration algorithm that is able to find deep connections in an short time and discriminate between them by looking at those that can be the most interesting. Our project aims to making the user know about the existing connections that can link music to a Place, by stimulating the curiosity of the user proposing non common relations.
\end{abstract}

\section{Introduction}
todo: change
Most of the available music recommenders suggest music without taking into consideration the user’s context, e.g., her mood, current location, activity. 
Finding music that suits a POI can be viewed as a context-aware recommendation problem, the place being the context for consuming the recommendation.         
The main challenge that one must face when addressing the above mentioned goal is related to the fact that POIs and music are two rather different domains, and there is no obvious way to match such heterogeneous items. However, with the advent of the Semantic Web, and specifically with its reference implementation in the Linked Data initiative, new opportunities arise to face the above difficulties.

\subsection{Motivation and context}
In the past, many tried to link POIs and music in many different way. xxx managed to link music and POIs by using a common set of tags that describe emotional properties both for music and POIs.
This method is completely supervised, in fact it is necessary to define in a clean way emotions for songs and POIs. This operation was done by hand and is hardly scalable.
Marius Kaminskas restrict the subspace of DBpedia by identifying classes belonging to the two domains of interest, and the relations existing between instances of such classes. In this way they managed to find relations in a set of constrained paths. Looking on the future works proposition, they proposed the idea to exploit arbitrary semantic relations between POIs and musicians, e.g. direct relations such as 'Gustav Mahler was the director of Vienna State Opera', and complex non-directed relations such as 'Ana Bel' (a famous Spanish singer) composed a song whose lyrics are about La Puerta de Alcala (a well-known POI in Madrid, Spain). 
In order to do this, they wanted to use Relfinder and set the connections' weights programatically. In fact one of the bad things was that connections weights were assigned by experts and were not provided in the paper.

What we intend to do is to find relations in a completely unsupervised way: this means that we intend to define a pathfinder that finds paths and a path discriminator that selects the interesting algorithms. Our intention is to limit to zero the need of human intervention in the hope that this approach will be completely data driven and easily scalable

\subsection{Research problems}
Finding relations inside DBpedia is a known problem which has been solved in the past. So we could simply look for POIs and Artists in Dbpedia and then link them using known algorithms.
Unfortunately DBpedia is not perfect: it is true that it contains POIs and Artists, but these entities are not cleanly defined, they usually contains dirty data that can tremendously affect our performances.
POIs and Artists can be found in highly specialized datasets, but unfortunately they do not contain links to eachother, and even worse they are not linked to DBpedia. This means that by using these databases alone we cannot find relations.
What is done in this cases is to link highly specialized datasets to DBpedia, in this way we can exploit the advantages of both worlds: highly specialized datasets give us confidence and trust in our main entities (POIs and Artists), while DBpedia allow us to find relations between them thanks to its generality.
The following two sections will explain in a detailed way the whole process of connecting these highly specialized datasets to DBpedia.

\subsection{Contributions}

\subsection{Report structure}
The report is divided into five sections, one for each part of the project. The first two parts are about linking external datasets to Dbpedia, this is also known as entity linking and we perform it in the optic of enriching our data with the inter-concept links present in DBpedia.
The third treats the pathfinder, a module of our system that is able to find deep connections between two entities in DBpedia.
The fourth describes the whole implementation of the REST API that allows the application to work smoothly.
The fifth part is about the mobile web application and the technologies involved in its functioning.

\section{Interlinking and enriching POIs}
\subsection{Problem description}
Finding relations inside DBpedia is a known problem which has been solved in the past. So we could simply look for POIs and Artists in Dbpedia and then link them using known algorithms.
Unfortunately DBpedia is not perfect: it is true that it contains POIs and Artists, but these entities are not cleanly defined, they usually contains dirty data that can tremendously affect our performances.
We start our search of connecting POIs and Artists from the POIs. POIs can be found in highly specialized datasets that contains clean data that we can trust, one of these datasets is 3cixty. Now, it would be optimal to find a direct mapping between a 3cixty POI and a DBpedia POI, in this way we could actually enrich the 3cixty POI with the context informations that can be found on DBpedia. Unfortunately this link does not exist and we must create it on our own by using an heuristic.

\subsection{Datasets: 3cixty, DBpedia, Wikidata}
\subsubsection{3cixty}
3cixty is a semantic dataset that contains aggregated informations of POIs and events. The dataset is the result of an aggregation process that join the POI informations from different datasets e.g. Facebook, Foursquare, Yelp \dots
In this project 3cixty is used as a trusted source of POIs for Nice.
POIs have many informations, such as the category, the address, comments from users, and coordinates. In order to get all the POIs in Nice, we simply used the POIs endpoint selecting the Nice city.
\subsubsection{DBpedia}
DBpedia is a project that aims at connecting the semantic world. We used it to get the POIs in Nice.
In order to get the POIs from DBpedia, it is necessary to perform SPARQL queries against the SPARQL endpoint. POIs are retrieved using two different sparql queries that target different kind of resources:
\begin{enumerate}
\item We perform a query that looks for all Geoentities that are in the area of nice. An entity is in the area of Nice if its coordinates are in a custom box that contains all the area of Nice.
Unfortunately there are many POIs that are without coordinates,so they are not selected with this query.
\item We perform some queries that get in a recursive way all entities that have as a parent the category of nice or nice and that are related to a place. The relation to a place is a list of classes that is done by hand and could be subjected to refinements.
\end{enumerate}
\subsubsection{Wikidata}
Wikidata is a project that aims at connecting the semantic world. We used it to get the POIs in Nice. Respect to DBpedia, Wikidata has a well defined ontology that is strictly enforced. This means that it is easier to select what we want. Moreover, most of POIs in wikidata have coordinates, so we can get them with a single query. Since wikidata supports complex geo-based queries, we made a query that looked for all POIs in a range of 25 Kilometers from the center of Nice.
\subsection{Entity Linking process}
Now that we have locally all our entities, we can link them. Having the entities locally is crucial, in fact it allows us to make way more complex operations respect to those that can be done with SPARQL queries.
\subsubsection{Properties}
The first phase is to actually get the interesting properties of the entities. Regarding those that come from DBpedia, we take use only the labels. We do not use coordinates because not all entities have them and we do not want to bias our algorithm using DBpedia coordinates, which is known that are not very accurate. Regarding the labels, we take all labels from all entities, in this way we are going to do a multilingual match because we know that sometimes this can make the difference.
Respect to Dbpedia, 3cixty entities have very accurate coordinates, and few labels. For this phase we do not need the coordinates, while regarding the labels, we keep the first one.
\subsubsection{Transformations}
At this point of the pipeline, we have labels that are encoded in unicode utf-8. Common similarity metrics accepts only plain ASCII strings as input, so we need to convert them into ASCII. Since utf-8 can represent way more characters than ASCII, we need to strip non ASCII characters from our labels. This is done in an intelligent way using the python library 'unidecode', that is able to convert from unicode to ASCII deleting only few characters. For example, an accented letter like 'è' can't be represented in ASCII. Most of the libraries delete all incompatible characters, while 'unidecode', when possible, converts it to the nearest ASCII character, in this case 'è' is converted to 'e'. In this way we can retain most of the original information after this conversion process.
The second step is probably the most important, and is the normalization of names. Some similarity metrics are extremely sensitive to noise. For example, when we use simple levenstein distance between ('aeroport','aeroport nice') is 5, while the distance ('aeroport','port') is 4. This is enough to raise some concerns: the simple addition of the city name can turn a perfect match into a mediocre one. In order to fix this problem we add the city name to all the labels where this is not present.
The third step is to strip all not alphabetical characters and to convert to lowercase all strings, this offcourse gives us cleaner strings to compare.

\subsubsection{Similarity measures and matching algorithm}
Now that strings are clean, we can proceed to the matching. A first problem rises: we want to match a 3cixty poi to a dbpedia poi, or we want to match a dbpedia poi to a 3cixty poi? In the first case we are sure that each 3cixty poi will be matched to a single dbpedia poi, but there is the risk that multiple 3cixty poi are matched to the same DBpedia poi. Alternatively, we match are sure that each DBpedia poi is matched to at most a 3cixty poi, but there is the risk that two DBpedia poi are matched to the same 3cixty poi.
Regarding DBpedia, we are sure that POIs are univoque, while we know that for the same POI, in 3cixty there are multiple ones.
Since at the end we will be using 3cixty coordinates, we opt for the first choice: if a DBpedia poi will be matched to multiple 3cixty pois, than we will follow an heuristic to choose the best match.
Regarding similarity metrics among strings, there are multiple ones, but most of them are based on levenstein distance.
Since each distance has its privileges and defects, we use multiple similarity measures, taken from the 'fuzzy wuzzy' python library.
For each 3cixty poi get the label and calculate the different similarity metric against all labels of all DBpedia POIs.

\subsubsection{Aggregations}
The aggregation algorithm can be described as follow:
\begin{enumerate}
\item For each similarity metric, take the three best matches and for them average the score of all similarity measures
\item The final match is the POI corresponding with the string with the highest score
\end{enumerate}
Why we don't simply take the average, or the minimum, or the maximum?
The reasoning behind this is that by taking the average we are flattening all the scores, and there is the risk that the string with the highest similarity is a string that is mediocre in all the different metrics. If a string is mediocre in all comparisons, then we do not trust it. Instead, with our methodology we are keeping only those matches that excel in some similarity metric, and then we take the one that performs better on average. We think that in this way we can filter out all those matches with average strings that matches almost anything.

\subsubsection{Filtering}
Now that the matching is done, we need to filter our results.
The first filter to do is to eliminate all those matches that match with Nice, but that have a label different from Nice. This usually happens because we added Nice to all strings. That addition was helpful, and it also created a safe default that now can be easily filtered.
Most of these match are acronym that match with Nice e.g. 'ATP Nice' matches with Nice because half of the string is Nice and is artificial, because we added it.

At this point we filter out all matches with a score below a given threshold.

Then, we have probably multiple 3cixty POIs connected to a single DBpedia POI.
In order to choose the most representative 3cixty POI, from whom we will take the coordinates, we have few possibilities:
\begin{enumerate}
\item K-medoids: We can actually cluster our points and look at the rapresentant of the biggest cluster 
\item Pick the POI which is nearest to the center
\item Pick the POI which is nearest to the DBpedia POI, if coordinates are available
\item Pick the POI which is nearest to the position found on OpenStreetMap
\item Link our data to GeoNames, a dataset of POIs
\end{enumerate}
The first method might be the best one, but it requires several precautions.
The second method is baiased towords pois in the center.
The third method can work only with POIs with coordinates, and gives good results if DBpedia coordinates are good, but we already know that this is not true.
The fourth method might be the best one, but we need to exit from the semantic world and hope that the API gives a good result, here the problem is that we have no control over the API.
The fourth method can work, but we would need to design another linking process.

For simplicity and reliability, we decided to use both OpenStreetMap API and Google Map API, in this way we have a much bigger confidence.

We make a request with the DBpedia label to the API, and we keep the POI which is nearest to one of them, in this case the minimum distance wins. If the distance is greater than a chosen threshold (500 meters), then the match is filtered out.

\subsection{Evaluation (Google Map / OpenStreetMap)}
The evaluation part is done using both the Google Map and OpenStreetMap. We calculated the Medium Square Error of the POIS to the positions recovered with the APIs, the error distance is the haversine distance.
\section{Interlinking and enriching Artists}
\subsection{Problem description}
\subsection{Datasets: DOREMUS, DBpedia, Wikidata}
\subsubsection{DOREMUS}
3cixty is a semantic dataset that contains aggregated informations of POIs and events. The dataset is the result of an aggregation process that join the POI informations from different datasets e.g. Facebook, Foursquare, Yelp \dots
In this project 3cixty is used as a trusted source of POIs for Nice.
POIs have many informations, such as the category, the address, comments from users, and coordinates. In order to get all the POIs in Nice, we simply used the POIs endpoint selecting the Nice city.
\subsubsection{DBpedia}
DBpedia is a project that aims at connecting the semantic world. We used it to get the POIs in Nice.
In order to get the POIs from DBpedia, it is necessary to perform SPARQL queries against the SPARQL endpoint. POIs are retrieved using two different sparql queries that target different kind of resources:
\begin{enumerate}
\item We perform a query that looks for all Geoentities that are in the area of nice. An entity is in the area of Nice if its coordinates are in a custom box that contains all the area of Nice.
Unfortunately there are many POIs that are without coordinates,so they are not selected with this query.
\item We perform some queries that get in a recursive way all entities that have as a parent the category of nice or nice and that are related to a place. The relation to a place is a list of classes that is done by hand and could be subjected to refinements.
\end{enumerate}
\subsubsection{Wikidata}
Wikidata is a project that aims at connecting the semantic world. We used it to get the POIs in Nice. Respect to DBpedia, Wikidata has a well defined ontology that is strictly enforced. This means that it is easier to select what we want. Moreover, most of POIs in wikidata have coordinates, so we can get them with a single query. Since wikidata supports complex geo-based queries, we made a query that looked for all POIs in a range of 25 Kilometers from the center of Nice.

\subsection{Entity Linking process (+ ElasticSearch)}
\subsubsection{Properties}
\subsubsection{Transformations}
\subsubsection{Similarity measures}
\subsubsection{Aggregations}
\subsubsection{Filtering}
\subsection{Evaluation}
todo: add source of isni
The evaluation of the found links is done by comparing the found links with those provided by ISNI. ISNI is the ISO certified global standard number for identifying the millions of contributors to creative works. By achieving these goals the ISNI will act as a bridge identifier across multiple domains and become a critical component in Linked Data and Semantic Web applications.

ISNI matches are made by humans, so we can consider them as a ground truth that can be trusted. By comparing our matches we want to see if our performances are acceptable and if we can discover new links between ISNI numbers and DBpedia.

todo: table with cool statistics
todo: graph image

We can state that the proposed method based on combinations of labels, classes and dates is able to match almost perfectly the artists of the two datasets. Moreover, we are able to find 2000 artists that at the time of writing are not present in the ISNI register.

\section{Path finder}
\subsection{Problem description}
todo: add citations
Once we have a direct mapping between entities that comes from specialized datasets and DBpedia, we actually need to link POIs and Artists.
The problem of finding relations between entities in DBpedia has already been approached and solved in the past, especially from Relfinder. The whole problem can be solved by writing a single query that finds a path of a given depth between two resources. This approach works well when datasets are relatively small and the two entities are two concepts that are semantically close. Instead, in our case we have concepts that are definetly far from each other, which implicates an high depth of an eventual path, and most importantly, we need to match hundreds of POIs with thousands of artists, This makes means that at the end the whole process will be repeated 100.000 times.
\subsection{Related work (Relfinder)}
The path finding problem has already been solved in the past with Relfinder: a tool that is able to find all paths with a specified depth between two entities. 
In fact, when we think at linking two entities, we need to set the position of the middle node and the number of middle nodes.

$$a-->b-->c-->d-->e=m$$
$$a-->b-->c-->d=m<--e$$
$$a-->b<--c=m-->d-->e$$

These are all different paths that links two entities 'a' and 'e'. In the simpler case, the arrows goes in one direction. In the second case they change direction one time.In the third case they change two times.
Relfinder manages to solve the problem by making different query that cover all possible combinations of the path specified above. As a constraint, in Relfinder the direction cannot change more than two times. There are still a huge amount of queries that explodes when the depth starts to be deep.
Here we intend as a deep depth a depth greater than 4, in fact this is like a magic threshold. 
Moreover, when we use Relfinder and set a greater depth we incur in two problems: 
\begin{enumerate}
\item The number of queries explodes because we need to consider all possible combinations
\item The time required by a single query explodes too, most of the times we cannot even do a query between two entities in a short time.
\end{enumerate}
In order to succeed we need to solve those two problem and we need to create a fast implementation, because at the end we will need to do 100.000 queries. 
\subsection{Proposed approach}
The problem is that DBpedia is small world and its average degree is 4.3, this means that on average we can reach every entity in at most 4-5 hops. Unfortunately our problem does not make part of that average, and many times we need 5-6 hops to connect two entities. The problem is that when we make a query that looks for paths with depth 6 we are basically querying the whole DBpedia, which at the time of writing is composed of 8.8 billion triples.
It is our belief that a query that tries to find a path between two entities, is converted by the SPARQL engine in what we can compare to a join for each hop. This, with the small world characteristic of DBpedia means that we are doing huge joins for each query.
Another way to visualize the query is a Breath First Search that starts from the source entity, and it expands hop after hop until the specified depth is reached. In this way is even more easy to see why we and up by querying billions of triples.
A common improvement of BFS is the two-way BFS, where the same algorithm is applied both at the source and the destination. In this way we can see that we can reach the same middle node by querying a way smaller portion of the graph.
Then, looking by hand at paths, we have seen that usually one direction inversion cover most of the cases, by limiting this we are actually simplifying a lot our problem. The, we have seen that for artists and POIs, the connection is most of the times the result of a common node. This means that the direction must change at least one time. We reduced all the different cases to one. In fact what we do is to make two separate queries with depth  from the source and from the destination. These queries return the list of all half-paths. We make a set of the involved nodes and then we intersect the two sets, in order to find the common nodes. At this point we operate joins that create full paths. The whole matching of POIs and Artists,100.000 combinations, requires only some hours against the weeks that would take relfinder.
\subsection{Selection Algorithm}
Now that we have all our path, we need to find the best ones, those that shows interesting links between POIs and Artists.
todo: lorenzo
\section{REST API}
The paths between a POI and the linked artists are exposed by a REST API that runs on Flask. Their extensible framework allows to develop several components requiring minimum boilerplate.
The main endpoint is the one that exposes the tracks and the paths related to a specific geoposition.
The whole process can be summarized in:
\begin{enumerate}
\item Take user position
\item Calculate nearest POIs and assign weights, from these create a univoque playlist name.
\item Select the tracks for each artist linked with the poi.
\item Select for each artist a compatible number of tracks to put in the playlist.
\item Create a playlist if it does not exist and add previously selected tracks.
\item Attach to each track the path between the poi and the track's artist.
\item Return the playlist ID and the track-path mappings as a JSON.
\end{enumerate}
\subsection{Spotify API usage}
Spotify provides a public API that allows to perform all the operations that can be done using the Spotify application. These comprend artist and tracks search, CRUD operations on user playlists and the listening of the tracks' preview.
\subsubsection{Authentication}
Since the end user listen to a spotify loaded dynamically by the application, there is no need for a client authentication. This is perfect because it allows to any user to use our application. Regarding the server, At startup it is performed a token based authentication, where the server get possess of a private token used to perform all CRUD operations on the playlist catalogue.
\subsubsection{Playlist}
A playlist can be dynamically created. At creation time, it is required to give a name to the playlist, and the whole playlist object, containing the univoque playlist ID is returned. Later it is possible to access to the playlist either using the playlist ID or using the playlist name. The playlist name is univoque inside the application by construction, while the playlist ID is univoque for all Spotify playlists.
\subsubsection{Artist}
An artist can be searched by name, this part is the most critical one because Spotify does not provide any automatic artist matching using the api. This means that specifying the full name of the artist is essential for the success of this operation.
This is done by using the Doremus person label property, which consists of name and surname of the artist. It is interesting to note that from this point of view DBpedia is not reliable because the person label can contains adjectives used for the disambiguation of this one, such as 'Composer'. If this adjectives are passed to the Spotify search endpoint, no artist is returned.
When a request is done, Spotify returns a list of matching artists. Each artist object contains different informations, among them the most useful are the genres and the popularity. 
In order to limit outliers, not only we select the most popular artist, but we check that it has at least 10 popular songs, where popular songs are special songs selected by spotify. The popularity metric algorithm is not public, but it is based on the playcounts and the recentness of the track.
Here we use the popularity as a criteria to select the correct artist, in fact we take always the most popular. The use of the genre specification would be interesting for an ulterior enhancement.
\subsubsection{Tracks}
Among all properties that caracterize a track, we only need the ID and the name.

\subsection{Playlist Endpoint}
\subsubsection{Nearest POIs selection algorithm}
In order to match a position with a list of POIs, the following algorithm is followed:
\begin{enumerate}
\item Calculate distance from the each POI to the user position. Since we are using coordinates, the distance is calculated using the haversine distance, which is very accurate and fast.
\item Add to the list the nearest POI, regardless of the distance.
\item Add the remaining two nearest POIs if their distance is inferior of 1 km.
\item For each element in the list, add 10 meters to the distance and then take the log2. By adding 10 meters we are adding a bias to all our distances, in this way positions too near to us are not over-weighted respect far distances. Then, the log2 squashes distances until 1 km to a restricted range of values.
\item Calculate repartition weights using the following equation:
$$w_i = \Bigg \lfloor 10 \dfrac{\dfrac{1}{\log_2 (d_i + 10)}}{\sum \dfrac{1}{\log_2 (d_i + 10)} } \Bigg \rceil$$
In this way we obtain normalized weights from 0 to 10 that are inversely proportional to the log base 2 of the distance. 
\end{enumerate}
At this point we have a direct mapping between POIs weights and a position
\subsection{Artist' tracks selection algorithm}
For each artist there is a direct mapping with its tracks.
The problem is: given a playlist name, find the correct tracks.
The playlist name encodes the weights of each poi, and each weight express exactly the number of tracks that must be chosen for the POI associated.
Since each POI is linked to different artists, we need a way to select songs from them in a balanced way.
What we do is to cycle between all the artists to take each time the most important song that is not already present in the playlist. In this way we obtain different songs from different artists, and since the artist and songs are static, the tracks in the playlist will be always the same. In the future we think to add randomicity to the whole selection process.

\subsection{Response generation}
\begin{enumerate}
\item Generate playlist name as described before
\item Select tracks as described before
\item Create a playlist if it does not exist and add previously selected tracks.
\item Attach to each track the path between the poi and the track's artist.
\item Return the playlist ID and the track-path mappings as a JSON.
\end{enumerate}

\section{Mobile Web Application}
\subsubsection{Music Player}
The music player is implemented using the Spotify Play Button. This is a special iframe that loads a simplified Spotify player for a specified playlist. This solution is extremely convenient because it simply needs a plylist ID to run. Since the playlists are created server side and are recovered using the REST API, the whole process is very simple.
Unfortunately it is not possible to manipulate programmatically the iframe and this limit the whole user experience. As an improvement, we could build a custom player that still uses the Spotify API.
\subsection{Google Map API}
\subsubsection{POIs placement}
todo: Lorenzo

\subsubsection{Navigation}

\subsection{Relation visualization (and Path format)}
todo: Lorenzo
\section{Conclusion and Future Work}

Your introduction goes here! Some examples of commonly used commands and features are listed below, to help you get started.

If you have a question, please use the support box in the bottom right of the screen to get in touch.

\section{Some \LaTeX{} Examples}
\label{sec:examples}

\subsection{Sections}

Use section and subsection commands to organize your document. \LaTeX{} handles all the formatting and numbering automatically. Use ref and label commands for cross-references.

\subsection{Tables and Figures}

Use the table and tabular commands for basic tables --- see Table~\ref{tab:widgets}, for example. You can upload a figure (JPEG, PNG or PDF) using the files menu. To include it in your document, use the includegraphics command as in the code for Figure~\ref{fig:frog} below.

% Commands to include a figure:
\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{frog.jpg}
\caption{\label{fig:frog}This is a figure caption.}
\end{figure}

\begin{table}
\centering
\begin{tabular}{l|r}
Item & Quantity \\\hline
Widgets & 42 \\
Gadgets & 13
\end{tabular}
\caption{\label{tab:widgets}An example table.}
\end{table}

\subsection{Mathematics}

\LaTeX{} is great at typesetting mathematics. Let $X_1, X_2, \ldots, X_n$ be a sequence of independent and identically distributed random variables with $\text{E}[X_i] = \mu$ and $\text{Var}[X_i] = \sigma^2 < \infty$, and let
$$S_n = \frac{X_1 + X_2 + \cdots + X_n}{n}
      = \frac{1}{n}\sum_{i}^{n} X_i$$
denote their mean. Then as $n$ approaches infinity, the random variables $\sqrt{n}(S_n - \mu)$ converge in distribution to a normal $\mathcal{N}(0, \sigma^2)$.

\subsection{Lists}

You can make lists with automatic numbering \dots

\begin{enumerate}
\item Like this,
\item and like this.
\end{enumerate}
\dots or bullet points \dots
\begin{itemize}
\item Like this,
\item and like this.
\end{itemize}

We hope you find write\LaTeX\ useful, and please let us know if you have any feedback using the help menu above.

\end{document}